---
title: "Practical 2"
author: "Jumping Rivers"
output:
  tufte::tufte_handout:
    highlight: tango 
    latex_engine: xelatex
    fig_width: 4
    fig_height: 4
  tufte::tufte_html: default
vignette: >
  %\VignetteIndexEntry{Practical 1}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r include = FALSE}
library(tufte)
knitr::opts_chunk$set(results = "hide", echo = FALSE, cache = FALSE)
```


```{r, echo = TRUE}
library("ggplot2")
```

\noindent Load the bond data set.

```{r, echo = TRUE, eval = FALSE}
bond = readRDS(file = "data/bond.rds")
```

```{r, echo = FALSE, eval = TRUE}
bond = readRDS(file = "../data/bond.rds")
```

## Scatter plots

Let's start with a basic scatter plot

```{r,fig.keep='none', echo = TRUE}
ggplot(data = bond, aes(x = Kills, y = Alcohol_Units)) +
  geom_point()
```

The arguments `x` and `y` are called aesthetics. For `geom_point`,
these parameters are required. This particular geom has other aesthetics:
`shape`, `colour`, `size` and `alpha`.^[These aesthetics are
  usually available for most geoms.] Experiment with other aesthetics. For example,

```{r,fig.keep='none', echo = TRUE}
ggplot(data = bond, aes(x = Kills, y = Alcohol_Units)) 
  geom_point(aes(colour = Actor))
```

\noindent or

```{r,fig.keep='none', tidy=FALSE, cache=TRUE, echo = TRUE}
ggplot(data = bond, aes(x = Kills, y = Alcohol_Units)) +
  geom_point(aes(colour = Actor,
                 shape = Nationality),
             size = 3)
```

**Remember** 

* If you are defining an aesthetic by a column of the data it needs to go **inside `aes()`**. E.g. `geom_point(aes(colour = Actor))`

* Otherwise, define it **outside** `aes()`. E.g. `geom_point(colour = "red", size = 2)`


Some aesthetics, like `shape` must be discrete.  If you try and use `shape = Relationships` you'll get an error. 
    
- What happens if you colour the points by `Relationships`?

```{r,fig.keep='none', tidy=FALSE, cache=TRUE, echo = TRUE}
ggplot(data = bond, aes(x = Kills, y = Alcohol_Units)) 
  geom_point(aes(colour = Relationships))
```
   
## Combining plots

The key idea with **ggplot2** is to think in terms of layers not in terms of plot "types". For example,

```{r,fig.keep='none', tidy=FALSE, echo = TRUE}
ggplot(bond, aes(x = Actor, y = Kills, colour = Nationality)) +
        geom_boxplot() + 
        geom_point(size = 3)
```

- What happens to the plot if you swap the order of the `geom_boxplot()`
  and `geom_point()` function calls?
  
- Notice that we defined the `aes(colour = Nationality)` in the `ggplot()` function. This means that the colouring is inherited by **both** `geom_boxplot()` and `geom_point()`. Try removing `colour` from `ggplot()` and defining just in `geom_boxplot(colour = "Nationality")`. What happens?

## Positions and flips

The bar geom has the following aesthetics: `x`, `colour`, `fill`,
`size`, `linetype`, `weight` and `alpha`. Here is a command to get
started:

```{r,fig.keep='none', echo = TRUE}
ggplot(bond, aes(x = Nationality)) + 
    geom_bar()
```

- As before, try different aesthetic combinations. 

```{r, fig.keep='none', echo = TRUE}
ggplot(bond, aes(x = Nationality)) + 
    geom_bar(aes(fill = Actor))
```

\noindent We can adjust the layout of this bar plot using ggplot's position
adjustments. The five possible adjustments are listed in table 1. The **default** adjustment is `stack`

```{r,fig.keep='none', tidy=FALSE, echo = TRUE}
ggplot(bond, aes(x = Nationality)) + 
    geom_bar(aes(fill = Actor), position = "stack")
```

- Try the other adjustments. 

```{r,echo=FALSE, fig.margin = TRUE, fig.cap="Barplot of ages using position = 'dodge'", out.width="\\textwidth", fig.height = 4}
ggplot(bond, aes(x = Nationality)) + 
    geom_bar(aes(fill = Actor), position = "dodge")
```

\begin{table}[t]
  \centering
  \begin{tabular}{@{}ll@{}}
    \toprule
    Adjustment & Description \\
    \midrule
    \texttt{dodge} & Adjust position by overlapping to the side \\
    \texttt{fill} & Stack overlapping elements; standardise stack height\\
    \texttt{identity} & Do nothing \\
    \texttt{jitter} & Jitter points \\
    \texttt{stack} & Stack overlapping elements \\
    \bottomrule
  \end{tabular}
  \caption{Position adjustments - table 4.5 in the ggplot2 book.}
  \label{T1}
\end{table}

\newpage

## Themes

Create **any** plot you like on either the `movies` or `bond` data set (or your own data if you like!) 

Experiment with the different theme options using `+ theme_NAME()`.
The in-built options are: 

* `theme_gray()` – signature ggplot2 theme
* `theme_bw()` – dark on light ggplot2 theme
* `theme_linedraw()` – uses black lines on white backgrounds only
* `theme_light()` – similar to linedraw() but with grey lines aswell
* `theme_dark()` – lines on a dark background instead of light
* `theme_minimal()` – no background annotations, minimal feel.
* `theme_classic()` – theme with no grid lines.
* `theme_void()` – empty theme with no elements

For example:

```{r movies, echo = FALSE}
movies = readRDS(file = "../data/movies.rds")
```

```{r themes, fig.keep='none', echo = TRUE}
ggplot(movies, aes(x = duration)) +
  
    geom_density(aes(fill = classification),
                 alpha = 0.5) + 
    theme_bw()
```




