---
title: "Practical 1"
author: "Jumping Rivers"
output:
  tufte::tufte_handout:
    highlight: tango 
    latex_engine: xelatex
    fig_width: 4
    fig_height: 4
  tufte::tufte_html: default
---

We will continue to investigate the movies data set.

* Make sure you've downloaded the GitHub repository

* Link: [bit.ly/nhs-ggplot2](bit.ly/nhs-ggplot2).

* Click on the .Rproj file to open the R project for today. This ensures that our working directory is correct!

* Open a new R script (`Ctrl+Shift+N` in RStudio.)

```{r, echo = TRUE, eval = FALSE}
library("ggplot2")
movies = readRDS(file = "data/movies.rds")
```

\noindent When loading in data, it's always a good idea to carry out a sanity check. I tend to use the commands

```{r,results='hide', echo = TRUE, eval = FALSE}
head(movies)
names(movies)
dim(movies)
View(movies)
```

## Scatter plots

(1) Make a basic scatter plot with votes on the x axis and rating on the y axis. Remember, we use the `ggplot()` function to create a basic plot and them `geom_point()` to add points.

```{r, eval = FALSE, echo = FALSE}
g = ggplot(movies, aes(x = votes, y = rating)) + 
  geom_point()
```

(1) Use the `labs()` function to change the axis labels & and title to something better

```{r, eval = FALSE, echo = FALSE}
g + labs(x = "No of Votes",
         y = "Average IMDB Rating", 
         title = "No of Votes vs Average IMDB Rating")
```

(1) The range of possible ratings is between 0 and 10, however because the maximum rating is below 10 the y-axis stops before 10. We can change the axis range using the `ylim` function. For instance, if our graph was stored in an object `g`, we could change the y-axis limit to (0,10) like so ` + ylim(0, 10)`.

Do this to your graph.

```{r, eval = FALSE, echo = FALSE}
g + ylim(0,10)
```


## Histograms

(1) Make a basic histogram of the year of releases. 

```{r, eval = FALSE, echo = FALSE}
ggplot(movies, aes(x = year)) +
  geom_histogram()
```

(1) Change the binwidth to one year using the `binwidth` argument in `geom_histogram()`

```{r, eval = FALSE, echo = FALSE}
ggplot(movies, aes(x = year)) + 
  geom_histogram(binwidth = 1)
```

(1)  Try changing `geom_histogram()` to `geom_density()`. You won't need a binwidth argument for the density plot. 

```{r, eval = FALSE, echo = FALSE}
ggplot(movies, aes(x = year)) + 
  geom_density()
```

# Bonus questions - an introduction to colours and fills

Let's go back to the classification bar chart we had in the slides.

```{r, eval = FALSE, echo = TRUE}
ggplot(movies, aes(x = classification)) + 
  geom_bar()
```

(1) Inside the `geom_bar()` function, try adding the argument `colour = "blue"`. What happens?

```{r, eval = FALSE, echo = TRUE}
ggplot(movies, aes(x = classification)) +
  geom_bar(colour = "blue")
```
<!-- It colours the outline blue-->

(1) Change `colour = "blue"` to `fill = "blue"`

<!-- It fills the bars in blue-->

(1) Experiment with other colours, you can find a list of colours that R takes using the `colours()` function i.e. run 
```{r, eval = FALSE, echo = TRUE}
colours()
```

(1) Remove `fill = "blue"` from `geom_bar()`. Now try adding `fill = classification` to the `aes()` function. What happens?

```{r, eval = FALSE, echo = TRUE}
ggplot(movies, aes(x = classification)) +
  geom_bar(aes(fill = classification))
```

<!-- It fills the bars in based on the variable classification-->